```{r}
library(ProjectTemplate); load.project()
```


# Gamma shift
```{r message=FALSE}
set.seed(1234)
Data <- rib[,c('user_id', 'block', 'rt')]
Data$user_id <- as.numeric(as.factor(Data$user_id))
names(Data) <- c('subject', 'trial', 'y')
N <- length(unique( Data$subject))

# Convert data to input format required by JAGS
jagsdata_overall <- list(subject=Data$subject, trial=Data$trial, y=Data$y, N=nrow(Data))
settings <- analysis_settings("quick")

m <- list()
models_overall <- list()
source('jags//jags_overall-2.r')
```


```{r}
# 3 power
cat(jags_overall_gamma(f='power3'), sep="\n")
models_overall$power3 <- run_jags(jags_overall_gamma(f='power3'),jagsdata_overall , c('alpha', 'beta'), showplots = FALSE)

samp <- models_overall$power3$samples
samp <- combine.mcmc(samp)
str(samp)

i <- i + 1
# JAGS uses shape and rate gamma parameterisation
plot(pgamma(1:100, samp[i, 'alpha[80]'], rate=samp[i, 'beta[80]']))


samp_a <- 

ijs <- 1:375


p <- NULL
for (ij in ijs) {
    alpha_var <- paste0('alpha[', ij, ']')
    beta_var <- paste0('beta[', ij, ']')

    p[ij] <-   mean(sapply(seq(nrow(samp)), function(X) pgamma(1, samp[X,alpha_var], rate=samp[X, beta_var])))
}

plot(density(p))
max(p)
mean(p)
mean(jagsdata_overall$y < 30)
min(rib$rt)
```

# constant error variance
```{r message=FALSE}
set.seed(1234)
settings <- analysis_settings("standard")
source('jags//jags_subtasks-2.r')
jagsdata <- get_data_subtasks()

script <- jags_subtasks_gamma(f="power3", theta2_constraint=FALSE,  gamma_constraint=FALSE, strategy_predictors=FALSE)
# sink("revisions/jags-subtasks-nostrategy-differenterror.jags")
# cat(script$script, sep='\n')
# sink()

# dput(script$parameters)
params <- c("theta1_1.mu", "theta1_2.mu", "theta1_3.mu", "theta1_1.sigma", 
    "theta1_2.sigma", "theta1_3.sigma", "theta2_1.mu", "theta2_2.mu", 
    "theta2_3.mu", "theta2_1.sigma", "theta2_2.sigma", "theta2_3.sigma", 
    "theta3_1.mu", "theta3_2.mu", "theta3_3.mu", "theta3_1.sigma", 
    "theta3_2.sigma", "theta3_3.sigma", 
    "sigma_1.mu", "sigma_1.sigma",
    "sigma_2.mu", "sigma_2.sigma",
    "sigma_3.mu", "sigma_3.sigma"
    )
fit <- run_jags("revisions//jags-subtasks-nostrategy-differenterror.jags", jagsdata, params, dic.run=TRUE)
get_dic(fit$dic)
```


# strategy parameterisation
```{r}
settings <- analysis_settings("standard")
script <- jags_subtasks_gamma(f="power3", theta2_constraint=FALSE,  
                              gamma_constraint=FALSE, strategy_predictors=TRUE)

sink("revisions/jags-subtasks-strategy.jags")
cat(script$script, sep='\n')
sink()

# dput(script$parameters)
params <- c("theta1_1.mu", "theta1_2.mu", "theta1_3.mu", "theta1_1.sigma", 
    "theta1_2.sigma", "theta1_3.sigma", "theta2_1.mu", "theta2_2.mu", 
    "theta2_3.mu", "theta2_1.sigma", "theta2_2.sigma", "theta2_3.sigma", 
    "theta3_1.mu", "theta3_2.mu", "theta3_3.mu", "theta3_1.sigma", 
    "theta3_2.sigma", "theta3_3.sigma", "lambda_1", "lambda_2", "sigma.mu", "sigma.sigma"
)
fit <- run_jags("revisions//jags-subtasks-strategy.jags", jagsdata, params, dic.run=TRUE)


```


# 10th percentile Quantiles, SD and Posterior predictive checks
```{r}
aggregate(ribs$rt, list(block=ribs$block, subtask=ribs$subtask), mean)
agg <- aggregate(ribs$rt, list(block=ribs$block, subtask=ribs$subtask), sd)
plot_stuff <- function(FUN, ...) {
    agg <- aggregate(rt~block+subtask, ribs, FUN)
    print(xyplot(rt~block, group=subtask, agg, type='l', auto.key=TRUE, ...))
}

plot_stuff(mean, ylab="mean rt")
plot_stuff(sd, ylab="sd rt")
plot_stuff(function(X) quantile(X, .10), ylab="10th percentile")
plot_stuff(function(X) quantile(X, .20), ylab="20th percentile")

agg <- aggregate(rt~block+subtask, ribs, sd)
agg <- aggregate(rt~subtask, agg, mean)

agg <- aggregate(rt~block+subtask, ribs, function(X) quantile(X, .1))
agg <- aggregate(rt~subtask, agg, mean)


```

# Posterior predictive checks
```{r}
```{r posterior_predictive_checks_version2_yhats}
source('jags/jags_subtasks-2.r')
source('jags/posterior-predictive-checks.r')


# prepare data file to be consistent with simulated datasets
Data <- ribs[,c('user_id', 'subtask', 'block', 'rt')]
names(Data) <- c('subject', 'subtask_text', 'trial', 'y')
Data$subtask <-  as.numeric(factor(Data$subtask_text, levels=c('informationgathering', 'filtering', 'timetableuse')))
Data$subject <- as.numeric(factor(Data$subject))

predictive_checks <- list()

# unconstrained without strategies
# 1. get posterior samples
# have a good burnin: 3000
# have many iterations: 10000 
# have thinning: 10
posterior_predictive_check_cus <- function(strategies=FALSE, constraints=FALSE, ...) {
    jags_model <- jags_subtasks_gamma(f='power3', strategy_predictors=strategies, theta2_constraint=constraints, 
                                      gamma_constraint=constraints, yhat=TRUE)
    pc_model <- run_jags(jags_model$script, jagsdata,  'yhat', dic.run=FALSE, showplots=FALSE)
    pc_samples <- do.call(rbind, pc_model$samples)
    
#     # 2. simulate from posterior
#     simulate_unconstrained_withoutstrateties <- function(thetas) {
#         simulate_data_subtasks_cus(as.list(thetas), constraints=constraints, strategies=strategies)$d
#     }
#     
    simulations <- lapply(seq(nrow(pc_samples)), function(X)
        data.frame(subject=jagsdata$subject, trial=jagsdata$trial, 
                                       subtask=jagsdata$subtask_index, y=pc_samples[X,]))    
    
    # 3. calculate statistics on simulated data and data
    x <- calculate_statistics(simulations, Data)
    summaryx <- summarise_statistics(x)
    list(summary_statistics=summaryx, raw_statistics=x)
}



settings <- list(n.chains= 1,  n.adapt= 200, burnin = 2000, n.iter= 10000, thin=1)
# settings <- list(n.chains= 1,  n.adapt= 200, burnin = 500, n.iter= 1000, thin=1)


predictive_checks$unconstrained_withoutstrategies <- posterior_predictive_check_cus(strategies=FALSE, constraints=FALSE)
predictive_checks$constrained_withoutstrategies <- posterior_predictive_check_cus(strategies=FALSE, constraints=TRUE)
predictive_checks$unconstrained_withstrategies <- posterior_predictive_check_cus(strategies=TRUE, constraints=FALSE)
predictive_checks$constrained_withstrategies <- posterior_predictive_check_cus(strategies=TRUE, constraints=TRUE)

lapply(predictive_checks, function(X) X$summary_statistics)

# create the summary table for posterior predictive checks 
extract_checks <- function(statistics_summary, statistics, digits=2) {
    combine_formatted <- function(x, digits) {
        formatted <- specify_decimal(x, digits)
        paste0(formatted['mean'], '\n', '(', formatted['low95'], ' - ', 
               formatted['high95'], ')',
               '\n', 'p=', formatted['pvalue'])
    }
    
    statistics_summary <- statistics_summary[,statistics]
    sapply(colnames(statistics_summary), 
           function(X) combine_formatted(statistics_summary[, X], digits))
}

predictive_checks_table <- sapply(predictive_checks, function(X) 
    extract_checks(X$summary_statistics, colnames(X$summary_statistics))
    )

dataset1 <- specify_decimal(predictive_checks[[1]]$summary_statistics['dataset', ], 2)
predictive_checks_table <- cbind(dataset1,  predictive_checks_table)
            
write.csv(predictive_checks_table, 'graphs/predictive_checks_table-revision-1.csv')

sink('graphs/predictive_checks_table-revision-1.csv', append=TRUE)
date()
settings
sink()

temp <- predictive_checks$constrained_withstrategies
```

# Deviance distributions
```{r}
# overall task
devdist <- list()
settings <- analysis_settings("publication")
devdist$power3 <- run_jags(jags_overall_gamma(f='power3'),jagsdata_overall , c('deviance'), showplots = FALSE, dic.run=FALSE)
devdist$exp3 <- run_jags(jags_overall_gamma(f='exp3'),jagsdata_overall , c('deviance'), showplots = FALSE, dic.run=FALSE)
samppower <- combine.mcmc(devdist$power3$samples)
sampexp <- combine.mcmc(devdist$exp3$samples)
library(psych)
dat <- data.frame(cbind(exp=as.numeric(sampexp), power=as.numeric(samppower)))
head(dat)
psych::describe(dat)

par(mfrow=c(2,1))
hist(sampexp, xlim=c(1850, 2000))
hist(samppower, xlim=c(1850, 2000))

settings <- analysis_settings("standard")
# subtask
model_grid <- expand.grid(f=c('power3', 'exp3'), theta2_constraint=c(TRUE, FALSE),  
                           gamma_constraint=c(TRUE, FALSE), strategy_predictors=c(TRUE, FALSE),
                           stringsAsFactors=FALSE)
model_grid$name <- paste(
    sapply(model_grid$f, switch, power3='p3', exp3='e3'),
    ifelse(model_grid$theta2_constraint, 'tc', 'tu'),
    ifelse(model_grid$gamma_constraint, 'gc', 'gu'),
    ifelse(model_grid$strategy_predictors, 'st', 'sf'), 
    sep="_")

model_fits <- list()
for (i in seq(nrow(model_grid))) {
    cat(i)
    script <- jags_subtasks_gamma(f=model_grid$f[i], 
                                  theta2_constraint=model_grid$theta2_constraint[i], 
                                gamma_constraint=model_grid$gamma_constraint[i], 
                                strategy_predictors=model_grid$strategy_predictors[i])
    model_fits[[ model_grid$name[i] ]] <- run_jags(script$script, jagsdata,  "deviance", dic.run=TRUE)
}

```

# exponential constrained no strategies (parameters)
```{r}
set.seed(1234)
settings <- analysis_settings("standard")
source('jags//jags_subtasks-2.r')
jagsdata <- get_data_subtasks()

script <- jags_subtasks_gamma(f="exp3", theta2_constraint=TRUE,
                              gamma_constraint=TRUE, 
                              strategy_predictors=FALSE)
fit <- run_jags(script$script, jagsdata,  script$parameters, dic.run=TRUE)

debugonce(create_parameter_table)
subtasks_ptable <- create_parameter_table(
    samples = list(
        e3_tc_gc_sf=fit$samples,
        e3_tc_gc_sf=fit$samples),
        dics = list(
            e3_tc_gc_sf=fit$dic,
            e3_tc_gc_sf=fit$dic),
            parameter_order =  c(          
                "theta1_1.mu", "theta1_2.mu", "theta1_3.mu",
           "theta2_1.mu", "theta2_2.mu", "theta2_3.mu",
           "theta3_1.mu", "theta3_2.mu", "theta3_3.mu",
          "gamma.mu", "theta2.mu","sigma.mu",  
          "theta1_1.sigma", "theta1_2.sigma", "theta1_3.sigma",
           "theta2_1.sigma", "theta2_2.sigma", "theta2_3.sigma",
           "theta3_1.sigma", "theta3_2.sigma", "theta3_3.sigma",
          "gamma.sigma", "theta2.sigma","sigma.sigma"))
write.csv(subtasks_ptable, 'graphs/subtasks_parameter_table-revision-1.csv', na="")


```

# priors
```{r}
# examination of cutting-off part of likelihood
x <- rnorm(1000000, 0, 1)
z_cutoff <- c(1, 2,3,4,5)
diffs <- sapply(z_cutoff, function(X) mean(x[x < X]))
cbind(z_cutoff, diffs)
```

```{r}
set.seed(1234)
Data <- rib[,c('user_id', 'block', 'rt')]
Data$user_id <- as.numeric(as.factor(Data$user_id))
names(Data) <- c('subject', 'trial', 'y')
N <- length(unique( Data$subject))

# Convert data to input format required by JAGS
jagsdata_overall <- list(subject=Data$subject, trial=Data$trial, y=Data$y, N=nrow(Data))
settings <- analysis_settings("standard")

m <- list()
models_overall <- list()
source('jags//jags_overall-2.r')

multiplier <- 10 # repeat for 2 and 10
models_overall$power3_2 <- run_jags(jags_overall_gamma(f='power3', prior_multiplier = multiplier),
                                  jagsdata_overall , c('theta1.mu', 'theta1.sigma', 
                                              'theta2.mu', 'theta2.sigma' ,
                                              'theta3.mu', 'theta3.sigma', 
                                               'sigma.mu', 'sigma.sigma'), dic.run=TRUE)

# 3 exponential
models_overall$exp3_2 <- run_jags(jags_overall_gamma(f='exp3', prior_multiplier=multiplier),
                                  jagsdata_overall , c('theta1.mu', 'theta1.sigma', 
                                              'theta2.mu', 'theta2.sigma' ,
                                              'theta3.mu', 'theta3.sigma', 
                                               'sigma.mu', 'sigma.sigma'), dic.run=TRUE)

get_dic(models_overall$power3_2$dic)
get_dic(models_overall$exp3_2$dic)

```


```{r}
set.seed(1234)
settings <- analysis_settings("standard")
source('jags//jags_subtasks-2.r')
jagsdata <- get_data_subtasks()

script <- jags_subtasks_gamma(f="power3", 
                              theta2_constraint=FALSE,  gamma_constraint=FALSE,  strategy_predictors=FALSE, 
                              prior_multiplier = 10)
fit <- run_jags(script$script, jagsdata,  script$parameters, dic.run=TRUE)


get_dic(fit$dic)

```


# Correlations between parameters
```{r}
source('jags//jags_overall-2.r')
settings <- analysis_settings("standard")

models_cor <- list()
models_cor$power3 <- run_jags(jags_overall_gamma(f='power3'),
                                  jagsdata_overall , c('theta1.mu', 'theta1.sigma', 
                                              'theta2.mu', 'theta2.sigma' ,
                                              'theta3.mu', 'theta3.sigma', 
                                               'sigma.mu', 'sigma.sigma'), dic.run=FALSE)
models_cor$exp3 <- run_jags(jags_overall_gamma(f='exp3'),
                                  jagsdata_overall , c('theta1.mu', 'theta1.sigma', 
                                              'theta2.mu', 'theta2.sigma' ,
                                              'theta3.mu', 'theta3.sigma', 
                                               'sigma.mu', 'sigma.sigma'), dic.run=FALSE)

cor_plot <- function(samp = models_cor$power3$samples) {
    samp <- data.frame(combine.mcmc(samp))
    print(round(cor(samp), 2))
    r <- cor(samp)
    print(mean(r[2,1],  r[4,3],  r[6,5],  r[8,7]))
    
    psych::pairs.panels(samp)
    round(print(c(t1m_t2m=r[3,5], t1m_t3m=r[3, 7], t2m_t3m=r[5, 7])), 2)
}
cor_plot(models_cor$exp3$samples)
cor_plot(models_cor$power3$samples)




```

```{r}
source('jags/jags_overall-2.r')
settings <- analysis_settings("standard")

returnfits_overall <- function(f='power3') {
    jags_object <- run_jags(jags_overall_gamma(f=f),jagsdata_overall , c('theta1.mu', 'theta1.sigma', 
                                              'theta2.mu', 'theta2.sigma' ,
                                              'theta3.mu', 'theta3.sigma', 
                                               'sigma.mu', 'sigma.sigma', 'yhat'), dic.run=FALSE, showplots=FALSE)
    fits_samples <- do.call(rbind, jags_object$samples)
    fits_samples <- fits_samples[,  grep('yhat', colnames(fits_samples), value=TRUE)]
    fits_means <- apply(fits_samples, 2, mean)
    list(fits_means=fits_means, jagsdata_overall=jagsdata_overall)
}
    
fits <- list() 
fits$power <- returnfits_overall('power3')
fits$exp <- returnfits_overall('exp3')

# reassemble the dataframe
yhatdata <- data.frame(subject=jagsdata_overall$subject, trial=jagsdata_overall$trial, 
                       yhat_power= fits$power$fits_means, 
                       yhat_exp= fits$exp$fits_means, y=jagsdata_overall$y)

yhat_group <- aggregate(yhatdata[, c('y', 'yhat_power', 'yhat_exp')], list(trial=yhatdata$trial), mean)

extract_model_overall_yhat <- function(yhat_var='yhat_power', variables=c("trial", "y"), data=yhat_group) {
    x <- data[,c(variables, yhat_var)]
    names(x)[names(x) == yhat_var] <- 'yhat'
    x$model <- yhat_var
    x
}
yhatdata_split <- lapply(grep('yhat', names(yhat_group), value=TRUE), function(X) extract_model_overall_yhat(X))
yhatdata_long <- do.call(rbind, yhatdata_split)
yhatdata_long$Model <- yhatdata_long$model
yhatdata_long$Model[yhatdata_long$model == "yhat_exp"] <- "Exponential"
yhatdata_long$Model[yhatdata_long$model == "yhat_power"] <- "Power"


# add error bars
std_error <- function(x) sd(x)/sqrt(length(x))
ses <- sapply(split(rib$rt,  rib$block), std_error)
ses <- data.frame(trial=as.numeric(names(ses)), se=ses)
yhatdata_long <- merge(  yhatdata_long, ses)




p <- ggplot(yhatdata_long, aes(trial, y), ) + geom_point() 
p <- p + geom_line(aes(x=trial, y=yhat, linetype=Model))
p <- p  + scale_y_continuous("Task completion time (sec)", limits=c(0, 50)) + 
    scale_x_continuous("Block", breaks=1:15) 
p <- p + geom_errorbar(aes(ymin=y-se, ymax=y+se), width=.1)
p

pdf(file="graphs/overallrtbyblockwithfits-revised.pdf", width=6, height=4)
p
dev.off()

```


# Posterior predictive checks (PPC)
```{r}
load('saved-fits/predictive_checks.RData')
str(predictive_checks$unconstrained_withoutstrategies)

ppc <- lapply(names(predictive_checks),  function(X) 
        unlist(predictive_checks[[X]]$raw_statistics$simulations['y1minus15_f_mean', ]))
names(ppc) <- names(predictive_checks)

ppcd <- rbind(data.frame(strategy=FALSE, constraints=FALSE, y=ppc$unconstrained_withoutstrategies),
      data.frame(strategy=FALSE, constraints=TRUE, y=ppc$constrained_withoutstrategies),
    data.frame(strategy=TRUE, constraints=FALSE, y=ppc$unconstrained_withstrategies),
    data.frame(strategy=TRUE, constraints=TRUE, y=ppc$constrained_withstrategies))
ppcd$strategy <- ifelse(ppcd$strategy, "Strategies", "No Strategies")
ppcd$constraints <- ifelse(ppcd$constraints, "Constrained", "Unconstrained")
ppc_data <- predictive_checks$unconstrained_withoutstrategies$raw_statistics$dataset$y1minus15_f_mean

# plotting
p <- ggplot(ppcd, aes(x=y)) +  geom_density()
p <- p + geom_vline(xintercept = ppc_data, linetype = 2)
p <- p + facet_grid(constraints~ strategy) 
p <- p + scale_x_continuous("Mean Filtering Improvement", limits=c(0, 8))
p <- p + scale_y_continuous("Density of Statistics Simulated from Posterior")
p

pdf('graphs/ppc-mean-filtering.pdf')
p
dev.off()
```


 